---
output: 
  html_document: 
    keep_md: yes
---
Reproducible Data - Project 1 
=============================

read in data

```{r echo = TRUE}
  data <- read.csv("activity/activity.csv")
  options(scipen=999, blogdown.method = "markdown")
```

Calculate total number of steps per day, ignore missing values. 

```{r echo = TRUE}
  steps_per_day <- aggregate(data$steps, by = list(data$date), FUN = sum, na.rm = TRUE)
  names(steps_per_day) <- c("date","steps")
```

Make histogram of total number of steps taken per day

```{r echo = TRUE}
  hist(steps_per_day$steps, xlab = "Steps", main = "Histogram of Steps per Day Oct/Nov 2012")
```

Calculate the mean and median of the steps taken per day

```{r echo = TRUE}
   mean_steps <- round(mean(steps_per_day$steps, na.rm = TRUE), digits = 0)
   median_steps <- median(steps_per_day$steps, na.rm = TRUE)
```

The mean number of steps per day is `r mean_steps`. The median is `r median_steps`.

Time series plot of the 5 minute interval and steps taken per interval

```{r echo = TRUE}
   steps_per_interval <- aggregate(data$steps, by = list(data$interval), FUN = mean, 
                                  na.rm = TRUE)
   names(steps_per_interval) <- c("interval", "steps")
   with(steps_per_interval, plot(interval, steps, type = "l", 
                                 main = "Average number of steps per interval"))
   max_steps <- max(steps_per_interval$steps)
   max_interval <- steps_per_interval$interval[steps_per_interval$steps == max_steps]
```

The interval with the maximum average number of steps is `r max_interval`.

Inputing missing values

Count number of rows with na

```{r echo = TRUE}
   rows_missing_step_data <- sum(as.numeric(is.na(data$steps)))
```

There are `r rows_missing_step_data` rows missing step data.

Imput the missing values. Use the previously calculated mean steps for the interval if a value is missing.

```{r echo = TRUE}
    
  imputed_data <- as.data.frame(data)

  for (i in 1:nrow(imputed_data)) {
  if (is.na(imputed_data[i,]$steps)) {
    get_interval <- data[i,]$interval
    imputed_data[i,]$steps <-   
              round(as.numeric(steps_per_interval$steps[steps_per_interval$interval  
                                                       == get_interval]), digits = 0)
  }
}


max(imputed_data$steps)
check_rows_missing_step_data <- sum(as.numeric(is.na(imputed_data$steps)))
   
   
```
There are `r check_rows_missing_step_data` rows missing data after imputing the data. 

Compute a histogram with imputed data. Compare to histogram without imputed data. 

```{r echo = TRUE}
  imputed_steps_per_day <- aggregate(imputed_data$steps, by =    
                                      list(imputed_data$date), FUN = sum)
  names(imputed_steps_per_day) <- c("date","steps")
  hist(imputed_steps_per_day$steps, xlab = "Steps", main = "Histogram of Steps per Day Oct/Nov 2012 (missing steps imputed")
  
   imputed_mean_steps <- round(mean(imputed_steps_per_day$steps), digits = 0)
   imputed_median_steps <- median(imputed_steps_per_day$steps)

```

The mean number of steps with imputing is `r imputed_mean_steps` and without imputing is `r mean_steps`.
The median number of steps with imputing is `r imputed_median_steps` and without imputing is `r median_steps`.

Differences in activity patterns between weekdays and weekends

```{r echo = TRUE}
    library(lattice)
    week_days <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
    imputed_data$date <- factor(ifelse(weekdays(as.Date(imputed_data$date)) %in% 
                                          week_days, "weekday",  "weekend"))
    agg_data <- aggregate(steps ~ interval + date, imputed_data, mean)
      
  
    
    xyplot(steps ~ interval | date, agg_data, type ="l", layout = c(1,2), xlab = 
              "Interval", ylab = "Number of steps")
    
```